---
title: "DDK"
author: "Jidde Jacobi"
date: "05/03/2021"
output: html_document
---
  
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library("rlang")
library("dplyr")
library("tidyr")
library("stringr")
library("ggplot2")
```

### Load data
```{r}
setwd("/Users/45598770/Documents/analysis/DDK")

ctrl <- read.csv("CTRLsheet.csv", encoding="UTF-8",stringsAsFactors = T)
pd <- read.csv("PDsheet.csv", encoding="UTF-8",stringsAsFactors = T)

ctrl <- cbind(group = "typical", ctrl)
pd <- cbind(group = "PD", pd)

df <- rbind(ctrl, pd)
rm(ctrl, pd)

# set variables to numeric
df <- df %>% 
 mutate_at(vars(!1:8), as.character) %>%
 mutate_at(vars(!1:8), as.numeric) # produces errors which is OK



# remove PD24 because of Parkinsonism diagnosis
df <- df[!df$subj == "PD 24 P",]
df$subj <- droplevels(df$subj)

# remove /pa/ trials
df <- df[!df$traj == "LA",]
df$traj <- droplevels(df$traj)
```

### Data wrangling

Extract info from filename, merge trialnumber and repetition
```{r}
df<- df %>%
  separate(fname, c("project", "prompt", "block", "remove", "remove2", "trialno", "repetition"), sep = "_") %>%
  mutate(trial=str_c(trialno,repetition)) %>%
  select(!c("project", "remove", "remove2", "trialno", "repetition", "outl", "yGONS",	"yPVEL",	"yNONS",	"yMAXC",	"yNOFFS",	"yPVEL2",	"yGOFFS", "vGONS",	"vPVEL",	"vNONS",	"vMAXC",	"vNOFFS",	"vPVEL2",	"vGOFFS"
))
```

Checks
```{r}
table(df$subj, df$seg)
table(df$seg, df$traj)
```

Remove outliers
```{r}
# df <- df[!(df$subj == "CTRL 14" & df$trial == "013201"),] # first reps extremely slow
df <- df[!df$sub == "CTRL 21",] # lengthens the vowel and has a weird prosodic structure
# df <- df[!df$sub == "CTRL 15",] # extremely slow
# df <- df[!(df$subj == "CTRL 11" & df$trial == "013201"),] # first trial, reps extremely slow
df <- df[!(df$subj == "PD 17" & df$trial == "013701"),] # misreading prompt
df <- df[!(df$subj == "PD 17" & df$trial == "014801"),] # subj laughing during trial
df <- df[!(df$subj == "PD 09" & df$trial == "013201"),] # task isn't clear to subj
df <- df[!(df$subj == "PD 14" & df$trial == "013801"),] # no clear gestures

# CTRL15, PD07, PD08, PD10 are extremely slow
```



```{r}
# gestural duration
df$dur <- df$tGOFFS-df$tGONS

# duration of acceleration phase
df$accphase <- df$tPVEL-df$tGONS

df <- df %>%
    group_by(subj, trial) %>%
    mutate(syldur = tGONS - lag(tGONS, default = first(tGONS))) %>%
    ungroup()

# duration of deceleration phase
df$decphase <- df$tMAXC-df$tPVEL

# count repetitions
df <- df %>% 
  group_by(subj, trial) %>% 
  mutate(rep = n()) %>%
    ungroup()

# count repetitions
df <- df %>% 
  group_by(subj, trial) %>% 
  mutate(rep = n()) %>%
  ungroup()

# calculate euclidean distances and time lag between successive repetitions
df <- df %>%
  group_by(subj, trial) %>%
  mutate(dist2previous = sqrt( (xMAXC - lag(xMAXC))^2 + (zMAXC - lag(zMAXC))^2 )) %>% # calculate euclidean distances
  mutate(temp.diff = tMAXC - lag(tMAXC)) %>% # calculate lag
  mutate(temp.stability = sd(temp.diff, na.rm = T)) %>%
  ungroup()


# calculate euclidean distances to centroid of point cloud, and variability within the cloud
df <- df %>%
  group_by(subj, trial) %>%
  mutate(xMean = mean(xMAXC), zMean = mean(zMAXC)) %>% # centroid of point cloud
  mutate(dist2centroid = sqrt((zMAXC - zMean)^2 + (xMAXC - xMean)^2 )) %>%
  mutate(spread = sd(dist2centroid)) %>%
  ungroup()

# sanity check centroid
p  <- ggplot(subset(df,  subj == "CTRL 01" & trial == "013701")) + geom_point(aes(x=xMAXC, y=zMAXC))
p + geom_point(aes(x=xMean, y=zMean), color="red", size=5)

p  <- ggplot(subset(df,  subj == "CTRL 06" & trial == "013501")) + geom_point(aes(x=xMAXC, y=zMAXC))
p + geom_point(aes(x=xMean, y=zMean), color="red", size=5)

p  <- ggplot(subset(df,  subj == "CTRL 07" & trial == "013501")) + geom_point(aes(x=xMAXC, y=zMAXC))
p + geom_point(aes(x=xMean, y=zMean), color="red", size=5)
df$dist2centroid[df$subj == "CTRL 07" & df$trial == "013501"]

p  <- ggplot(subset(df,  subj == "CTRL 07" & trial == "013601")) + geom_point(aes(x=xMAXC, y=zMAXC))
p + geom_point(aes(x=xMean, y=zMean), color="red", size=5)
rm(p)
```


Set order of columns
```{r}
df <- df %>%
  select(subj, group, block, trial, prompt, seg, traj, note, repetitions, dur,accphase, decphase, rep, dist2previous, dist2centroid, temp.diff, everything())
```


Retrieve syllable rate

```{r}
# retrieve number of rep per trial
df <- df %>%
  group_by(subj,trial) %>%
  mutate(no.reps = n())


# remove < 2 repetition trials
df <- df[df$no.reps > 2,]

# select only first and last rep
df_lim <-df %>%
  group_by(subj,trial) %>%
  slice(c(1, n())) %>%
  mutate(whichRep = if_else(row_number() == 1, 'first', 'last') ) %>%
  ungroup()

# retrieve GONS of first rep
start <- df_lim %>%
  group_by(subj,trial) %>%
  filter(whichRep == "first") %>%
  rename(start = tGONS) %>%
  select(subj,trial,start) %>%
  ungroup()

# retrieve GOFFS of final rep and merge with GONS
rep.dur <- df_lim %>%
  group_by(subj,trial) %>%
  filter(whichRep == "last") %>%
  rename(end = tGOFFS) %>%
  select(subj,trial,end) %>%
  inner_join(start, by=c("subj","trial")) %>%
  select(subj, trial, start, end) %>%
  ungroup()
  
# merge
df <- df %>%
 right_join(rep.dur, by = c("subj","trial")) %>%
  group_by(subj,trial) %>%
  mutate(mean.rep.dur = ((end - start)/n())) %>%
  ungroup()

rm(df_lim, rep.dur, start)

```



### Plots
Examine speech rate by looking at mean duration of a repetition in every trial
```{r}
ggplot(df[df$block == "DDK1" & df$prompt == "TA",], aes(x=subj, y=mean.rep.dur, color=group)) + geom_boxplot() + labs(title="mean duration /ta/ repetition block 1", y="duration (ms)", x="Subject") 
ggplot(df[df$block == "DDK2" & df$prompt == "TA",], aes(x=subj, y=mean.rep.dur, color=group)) + geom_boxplot() + labs(title="mean duration /ta/ repetition block 2", y="duration (ms)", x="Subject") 
```


Examine speech rate by looking at mean duration of a repetition in every trial
```{r}
ggplot(df[df$block == "DDK1" & df$prompt == "KA",], aes(x=subj, y=mean.rep.dur, color=group)) + geom_boxplot() + labs(title="mean duration /ka/ repetition block 1", y="duration (ms)", x="Subject") 
ggplot(df[df$block == "DDK2" & df$prompt == "KA",], aes(x=subj, y=mean.rep.dur, color=group)) + geom_boxplot() + labs(title="mean duration /ka/ repetition block 2", y="duration (ms)", x="Subject") 
```

Compare variation around centroid for /ta/
```{r}
ggplot(df[df$block == "DDK1" & df$prompt == "TA",], aes(x=subj, y=dist2centroid, color=group)) + geom_boxplot() + labs(title="Distance to center of cloud in /ta/", y="Euclidian distance (mm)", x="Subject") 
ggplot(df[df$block == "DDK2" & df$prompt == "TA",], aes(x=subj, y=dist2centroid, color=group)) + geom_boxplot() + labs(title="Distance to center of cloud in /ta/", y="Euclidian distance (mm)", x="Subject") 
```


Compare variation around centroid for /ka/
```{r}
ggplot(df[df$block == "DDK1" & df$prompt == "KA",], aes(x=subj, y=dist2centroid, color=group)) + geom_boxplot() + labs(title="Distance to center of cloud in /ka/", y="Euclidian distance (mm)", x="Subject") 
ggplot(df[df$block == "DDK2" & df$prompt == "KA",], aes(x=subj, y=dist2centroid, color=group)) + geom_boxplot() + labs(title="Distance to center of cloud in /ka/", y="Euclidian distance (mm)", x="Subject") 
```


Compare displacement between successive repetitions for /ta/
```{r}
ggplot(df[df$block == "DDK1" & df$prompt == "TA",], aes(x=subj, y=dist2previous, color=group)) + geom_boxplot() + labs(title="Distance between successive repetitions in /ta/", y="Euclidian distance (mm)", x="Subject") 
ggplot(df[df$block == "DDK2" & df$prompt == "TA",], aes(x=subj, y=dist2previous, color=group)) + geom_boxplot() + labs(title="Distance between successive repetitions in /ta/", y="Euclidian distance (mm)", x="Subject") 
```


Compare displacement between successive repetitions for /ka/
```{r}
ggplot(df[df$block == "DDK1" & df$prompt == "KA",], aes(x=subj, y=dist2previous, color=group)) + geom_boxplot() + labs(title="Distance between successive repetitions in in /ka/", y="Euclidian distance (mm)", x="Subject") 
ggplot(df[df$block == "DDK2" & df$prompt == "KA",], aes(x=subj, y=dist2previous, color=group)) + geom_boxplot() + labs(title="Distance between successive repetitions in in /ka/", y="Euclidian distance (mm)", x="Subject") 
```


Temporal stability
```{r}
ggplot(df[df$block == "DDK1",], aes(x=subj, y=temp.stability, color=group)) + geom_boxplot() + labs(title="Temporal stability block 1", y="Variability", x="Subject")

ggplot(df[df$block == "DDK2",], aes(x=subj, y=temp.stability, color=group)) + geom_boxplot() + labs(title="Temporal stability block 2", y="Variability", x="Subject")
```



Across syllables
```{r}
#boxplots
ggplot(df[df$block == "DDK1",], aes(x=subj, y=mean.rep.dur, color=group)) + geom_boxplot() + labs(title="Gestural duration block 1", y="Gestural duration (ms)", x="Subjects") + ylim(0, 750)
ggplot(df[df$block == "DDK2",], aes(x=subj, y=mean.rep.dur, color=group)) + geom_boxplot() + labs(title="Gestural duration block 2", y="Gestural duration (ms)", x="Subjects") + ylim(0, 750)
```
Per syllable
```{r}
#boxplots
ggplot(df[(df$block == "DDK1" & df$prompt == "PA"),], aes(x=subj, y=mean.rep.dur, color=group)) + geom_boxplot() + labs(title="Gestural duration /p/ block 1", y="Gestural duration (ms)", x="Subjects") + ylim(0, 750)
ggplot(df[(df$block == "DDK2" & df$prompt == "PA"),], aes(x=subj, y=mean.rep.dur, color=group)) + geom_boxplot() + labs(title="Gestural duration /p/ block 2", y="Gestural duration (ms)", x="Subjects") + ylim(0, 750)
```

```{r}
#boxplots
ggplot(df[(df$block == "DDK1" & df$prompt == "TA"),], aes(x=subj, y=mean.rep.dur, color=group)) + geom_boxplot() + labs(title="Gestural duration /t/ block 1", y="Gestural duration (ms)", x="Subjects") + ylim(0, 750)
ggplot(df[(df$block == "DDK2" & df$prompt == "TA"),], aes(x=subj, y=mean.rep.dur, color=group)) + geom_boxplot() + labs(title="Gestural duration /t/ block 2", y="Gestural duration (ms)", x="Subjects") + ylim(0, 750)
```

```{r}
#boxplots
ggplot(df[(df$block == "DDK1" & df$prompt == "KA"),], aes(x=subj, y=mean.rep.dur, color=group)) + geom_boxplot() + labs(title="Gestural duration /k/ block 1", y="Gestural duration (ms)", x="Subjects") + ylim(0, 750)
ggplot(df[(df$block == "DDK2" & df$prompt == "KA"),], aes(x=subj, y=mean.rep.dur, color=group)) + geom_boxplot() + labs(title="Gestural duration /k/ block 2", y="Gestural duration (ms)", x="Subjects") + ylim(0, 750)
```
